---
title: "Biology Enrollment Trends"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

Set things up
```{r}
rm(list = ls())
library(here)
library(tidyverse)
```
Pull in data
```{r}
DF<-read.csv(here("Data", "CB_enroll_f17_s22.csv"), stringsAsFactors = T)
```

Rename colums and set things up a bit
```{r}
DF<-DF %>% rename(Course = Course.Name)
```
Let's add a column for majors/non-majors
```{r}
NM<-which(DF$CB_type == "NM")
DF$major <- "Yes"
DF$major[NM]<- "No"
```

Let's get a sense of open seats

```{r}
DF<-DF %>% mutate(
  open_seats = Cap - Enrollment)

```

Now just get fall data

```{r}
Fall<-DF %>% filter(Semester == "fall")
```


Now make a spagetti plot

```{r}
ggplot(Fall, aes(Year, Enrollment, color = Course, group = Course))+
  geom_point()+
  geom_line()
```

And now do again but omit GB
```{r}
fall_noGB <-Fall %>% filter(Course_num_official != "BIOL-101")
```

```{r}
ggplot(fall_noGB, aes(Year, Enrollment, color = Course, group = Course))+
  geom_point()+
  geom_line()+
  facet_wrap(~CB_type)
```
How are we for percent capacity?

```{r}
ggplot(fall_noGB, aes(Year, Percent_cap, color = Course, group = Course))+
  geom_point()+
  geom_line()+
  facet_wrap(~CB_type)+
  geom_hline(yintercept = 100, color = "black")
```


Now get some summary data, including
Number of courses, by con bio type,  year, and semester
Number of enrolled seats by con bio type, year, and semester
Number of available seats by con bio type, year, and semester
```{r}
fall_by_type<-fall_noGB %>% group_by(Year, CB_type) %>% summarize(
  num_classes = length(unique(Course)),
  hours = sum(Hours),
  FTE = hours/9,
  avg_enroll = mean(Enrollment, na.rm = T),
  sd_enroll = sd(Enrollment, na.rm = T),
  enrolled_seats = sum(Enrollment, na.rm = T),
  available_seats = sum(open_seats, na.rm = T),
  percent_cap = (enrolled_seats/(enrolled_seats+available_seats))*100
)
```
And just across fall semesters in general
```{r}
fall_averages<-fall_noGB %>% group_by(Year, major) %>% summarize(
  num_classes = length(unique(Course_num_grouped)),
  hours = sum(Hours),
  FTE = hours/9,
  avg_enroll = mean(Enrollment, na.rm = T),
  sd_enroll = sd(Enrollment, na.rm = T),
  enrolled_seats = sum(Enrollment, na.rm = T),
  available_seats = sum(open_seats, na.rm = T),
  percent_cap = (enrolled_seats/(available_seats+enrolled_seats))*100
)
```
Let's look at some plots
```{r}
ggplot(fall_averages, aes(Year, percent_cap, group = major, color = major))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept = 100, linetype = "dashed", color = "red")
```
```{r}
ggplot(fall_averages, aes(Year,enrolled_seats, group = major, color = major))+
  geom_point()+
  geom_line()+
  geom_point(aes(Year, available_seats), size = 1.5)+
  geom_line(aes(Year, available_seats), size = 1.5)
```
I'd like to pivot_longer so I can have a category of enrolled vs available seats. Do this for fall_averages ad fall_by_type

```{r}
fall_averages<-pivot_longer(fall_averages, enrolled_seats:available_seats, names_to = "seat_type", values_to = "Num_seats")

fall_by_type<-pivot_longer(fall_by_type, enrolled_seats:available_seats, names_to = "seat_type", values_to = "Num_seats")
```
Now plot
```{r}
ggplot(fall_by_type, aes(Year, Num_seats, color = seat_type, group = seat_type))+
  geom_point()+
  facet_wrap(~CB_type)
```








